
# Installation and Preparation {#installation}

## Installing via CRAN

You can install eyeTrackR via **CRAN** in the same way that you would install most R packages - in RStudio, head over to the **Packages** tab, and hit **Install**. Then search for eyeTrackR.

## Installing via Devtools

From R using the **devtools** package:

```{r, eval=F}
library(devtools)
install_github('hjgodwin/eyetrackR/source')
```

## Files you'll Need

To start with, you'll need to output three different types of report from [SR Research DataViewer](https://www.sr-research.com/data-viewer/). You'll need:

* Fixation Report
* Message Report
* Interest Area Report

When I generate these reports, I typically get DataViewer to output every single column of data it can. It has a whole bunch of different ones, but my preference is to be greedy. The simple reason for this is that DataViewer can take a long, long time to load up and output all your data, so you don't want to have to go back into it and do everything again just because you've missed a vital column. That being said, if you have a lower-spec or older computer, or are short on space, you may struggle with these files. 

When saving these reports, it's best to save them as tab-delimeted text files (.txt).

## R Skills you'll Need

Before using eyeTrackR and the output it provides, you may well want to get familiar with the excellent data.table package. I use that extensively in this guide, so you may want to learn about that first, or in addition to, going through this.

There's a very helpful introduction to using data.table available [here.](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html)

## Removing eyeTrackR

Please don't!!

Ok, if you want to, either to update it or because you just want to get rid of it, run the following from R:

```{r, eval=F}
remove.packages(pkgs="eyeTrackR")
```




